---
title: "mapping-cheat-sheet"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mapping-cheat-sheet}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(OSButils)
```



#' Generate Postal Area Stats
#'
#' @description When perfoming mapping tasks, it is necessary to have several
#'   pieces of information about the spacial level you are working at. You will
#'   want to know the more detailed description of the area, and the population
#'   in that area.
#'
#' @return
#' @export
#'
#' @examples
mmd_ImpGeneratePostalAreaStats <- function(){
        pop_data <- readr::read_csv("post area population data")

        # Postcode area names and volumes into one table pop_data - copied from Wikipedia
        pa_names <- readr::read_csv("pa names file") %>%
                dplyr::select(
                        PostcodeArea = `Postcode area`,
                        pa_name = `Postcode area name[1][3]`
                ) %>%
                # Remove footnote indicators i.e. [1] from the names
                dplyr::mutate(pa_name = gsub("\\[[0-9]\\]", "", pa_name))

        pop_data <-
                dplyr::left_join(pop_data,
                          pa_names,
                          by = "PostcodeArea"
                )

        pop_data
}

#' Process the penetration values by Postal Area
#'
#' @description A wrapper for the process of determining penetration by population
#'   and then breaking that down into percentiles.
#'
#' @param input_ds the input dataset
#' @param reference_areas The reference data for the UK postal area stats
#' @param .target_var The target variable to compute densities
#'
#' @return
#' @export
#'
#' @examples
mmd_PurProcessTargetByArea <- function(input_ds, reference_areas = mmd_ImpGeneratePostalAreaStats(),
                                       .target_var){

        input_ds %>%
                dplyr::right_join(
                        reference_areas
                        ,    by = c("Postcode.Area" = "PostcodeArea")
                ) %>%

                # many missing postal sectors without leads so make them zero
                dplyr::mutate_at(dplyr::vars({{.target_var}}), tidyr::replace_na, 0) %>%
                dplyr::filter(.data$Postcode.Area != "ZE") %>%

                # create pen which is penetration of leads per thousand
                dplyr::mutate(pen = {{.target_var}} / (.data$Population/1000)) %>%



                # create pencentiles from 0 worst to 1 best
                # pen_percentile_leads is the percentile for penetration of leads (takes into account the population
                # percentile_leads = percentile for number of leads
                dplyr::mutate(pen_percentile = ifelse({{.target_var}} == 0, 0 , column_percentiles(.data$pen)  ))
}

#' Getting Postcode area geometries
#'
#' @description When perfoming mapping tasks, it is necessary to have several
#'   pieces of information about the spacial level you are working at. The most
#'   important is the geometries of the region, this is the data which can
#'   instruct the mapping program to draw the regions.
#'
#'   We also simplify the geometries in the following ways:
#'   - Change the projection of the boundary data so that it can be plotted by
#'   leaflet
#'
#'   - Simplify using the mapshaper library, in order to produce a lower
#'   resolution map with a much smaller file size. The default behaviour
#'   is to keep 5% of the points that make up the boundaries. This seemed
#'   to work well in this case.
#'
#' @return
#' @export
#'
#' @examples
mmd_ImpGenerateGeometries <- function(){

        sf::st_read("J:/OCCAM/OCCAM.806/MapInfo_Data/Postal", layer = "POSTAREA") %>%
                sf::st_transform( "+init=epsg:4326") %>%

                # Simplify boundary data
                rmapshaper::ms_simplify() %>%
                dplyr::rename(Postcode.Area = .data$PostArea) %>%
                dplyr::mutate_at(dplyr::vars(.data$Postcode.Area), as.character) %>%
                dplyr::select(-Standard_Region)
}


# perc_pal_ace <- colorBin("Greens", domain = c(0, 1), bins = 5)

#' Percent Label Formatter
#'
#' @description A Wrapper for the percent formatting.
#'
#' @param ... The generic grouping variables, allowing for multiple aggregates
#' @param percent Flag for if we want to format the string as a percent or not
#'
#' @return
#' @export
#'
#' @examples
percentLabelFormat <- function(..., percent = TRUE) {
        if (percent) {
                function(type = "bin", cuts, ...) {
                        scales::percent(cuts)
                }
        } else {
                leaflet::labelFormat(...)
        }
}

#' Percentiles Generation
#'
#' @description Function to convert a vector into its percentile
#'   scores.
#'
#' @param column The targeted vector
#'
#' @return
#' @export
#'
#' @examples
column_percentiles <- function(column) {
        stats::ecdf(column)(column)
}



#' Extract the postcode Ares
#'
#' @description We commonly need to extract the postcode area from
#'   a postcode, this function wraps the regex extraction process.
#'
#' @param str_x The string containing the postcode.
#'
#' @return
#' @export
#'
#' @examples
pfl_PurExtractPostcodeArea <- function(str_x) {
        stringr::str_match(str_x, "^([a-zA-Z]{1,2})\\d{1,2}")[2]
}


#' Wrapper for the generation of the maps
#'
#' @param input_ds the input dataset
#' @param brand_name the brand,  which the data should be restricted
#'   to
#'
#' @return
#' @export
#' @importFrom rlang .data
#'
#' @examples
mmd_PurWrapMappingGen <- function(input_ds, brand_name){
        input_ds %>%
                dplyr::filter(.data$Brand == brand_name) %>%
                dplyr::pull(.data$Geom.Data) %>%
                {.[[1]]} %>%

                leaflet::leaflet() %>%
                # sales per 000 percentile
                leaflet::addRectangles(lat1 = 49, lat2 = 60, lng1 = -9.5, lng2 = 3, opacity = 1, color =  "#fff", fillOpacity = 1) %>%
                leaflet::addPolygons(fillOpacity = 1, color = "black", weight = 1, label = mmd_PurWrapLabelsProc(input_ds = input_ds, brand_name = brand_name), fillColor = ~perc_pal_ace(.data$pen_percentile), group = paste0(brand_name," Purchases per 000 percentile")) %>%
                leaflet::addLegend(title = paste0(brand_name," Ranking"), opacity = 0.8, position = "bottomright", pal = perc_pal_ace, group = "Percentile Legend", values = ~pen_percentile, labFormat = percentLabelFormat())
}
